---
layout: post
title: "Cardio Good Fitness Analysis "
author: "Spurthi"
date: 10/07/2020
categories: r , Exploratory Data Analysis
output:
  html_document:
    toc: yes 
    toc_depth: 4
---

## 1. Project Objective
  
  The objective of this report is to explore the Cardio Good Fitness data set in 
R and extract insights about the data.

This exploration report will consist of the following:
  
* Importing the dataset in R 
* Understanding the structure of dataset
* Graphical Exploration
* Descriptive Statistics
* Generate insights and recommendations that will help the company in targeting 
new customers

## 2. Assumptions 

The dataset provided is a sample of the population for customers that bought the 
treadmill product of Cardio Good Fitness Store. 

## 3. Exploratory Data Analysis - Step by step approach

```{r echo = FALSE}
#======================================================================= 
# 
# Exploratory Data Analysis - CardioGoodFitness 
# 
#=======================================================================
```

### 3.1 Environment Set up and Data Import 

#### 3.1.1 Install necessary packages and load libraries
```{r,warning=FALSE, message=FALSE}


library(ggplot2)# To create plots
library(gridExtra)# To plot multiple ggplot graphs in a grid
library(dplyr) # For data manipulation
library(knitr)# Necessary to generate sourcecodes from a .Rmd File 
library(corrplot) # To plot correlation plot between numerical variables
library(reshape2)# To easily transform data between wide and long formats
library(plotly) # R package for creating interactive web-based graphs 
library(ggbeeswarm) #R package for plotting categorical scatter plots
library(scales) # R package for internal scaling infrastructure used by ggplot2 
library(webshot)

```

#### 3.1.2 Set up working Directory

```{r}
# Setup Working Directory 
# Rmd considers it's directory as the working directory.
# If the Rmd and data file is in the same directory there is no need of 
# setting up the working directory.

# While working with .R files,we need to set up the working directory using the following command
# setwd("C:/Users")

```


#### 3.1.3 Import and Read the Dataset

```{r}
# Read Input File 
Fitness = read.csv("CardioGoodFitness.csv")
```


#### 3.1.4 Global options settings
```{r}
# Global options settings 
options(scipen=999) # turn off scientific notation like 1e+06
```

### 3.2 Variable Identification

#### 3.2.1 Variable Insights

**1. Checking the dimension of the dataset**
```{r}
dim(Fitness)

```
* The dataset has 180 rows and 9 columns


**2. Checking the structure of the dataset**
```{r}
str(Fitness)

```


**3. Checking top 10 records**
```{r}
head(Fitness, 10)

```
* Product column shows the treadmill code
* Age column shows the age of customer
* Gender - The data is consistent
* Education - No Of Years of Education the customer received
* MaritalStatus - Shows the single or married marriage status
* Usage - Shows how often the customer wants to use the treadmill
* Fitness - Shows level of Fitness according to the customer
* Income - Shows the income of the customer
* Miles - Shows the number of miles the customer wants to run/walk on the 
  treadmill.


**4. Checking bottom 10 records**
```{r}
tail(Fitness, 10)

```
* Data is consistent with the top, no missing data or unknowns.


**5. Checking the summary set of the data** 
```{r}
summary(Fitness)

```


### 3.3 Univariate Analysis

**1. Observation based on gender**
```{r}
## Bar Chart for Customers By Gender
Customers_By_Gender_Count = Fitness %>% count(Gender)

ggplot(Customers_By_Gender_Count, 
       aes(x = Gender, 
           y = n)) + 
  geom_bar(fill = "light blue", color="white",stat = "identity") +
  geom_text(aes(label = n), 
            vjust=-0.5) +
  labs(x = "Gender", 
       y = "Number of Customers", 
       title  = "Number of Customers By Gender")
```

* Checking the number of customers by gender.
* We have more number of male customers vs female customers.


**2. Observation based on Product**
```{r}
## Bar Chart for Customers By Product
ggplot(Fitness, aes(x = Product)) + 
  geom_bar(fill = "light blue", color="white") +
  labs(x = "",
       y = "Number of Customers",
       title = "Number of Customers By Product") +
  coord_flip()

```


* TM195 seems to be the most popular treadmill product.

```{r}
# Function to draw histogram and boxplot of numerical variables using ggplot


plot_histogram_n_boxplot = function(variable, variableNameString, binw){
  h = ggplot(data = Fitness, aes(x= variable))+
    labs(x = variableNameString,y ='No Of Custs')+
    geom_histogram(fill = 'light blue',col = 'white',binwidth = binw)+
    geom_vline(aes(xintercept=mean(variable)),
               color="black", linetype="dashed", size=0.5)
  b = ggplot(data = Fitness, aes('',variable))+ 
    geom_boxplot(outlier.colour = 'blue',col = 'blue',outlier.shape = 19)+
    labs(x = '',y = variableNameString)+ coord_flip()
  grid.arrange(h,b,ncol = 2)
}


```

**3. Observations on Age**

```{r}
plot_histogram_n_boxplot(Fitness$Age,'Age',10)
```

* The distribution of age is right skewed.
* Majority of customers that purchased the thread mills are in their 20's

**4. Observations on Education**

```{r}

plot_histogram_n_boxplot(Fitness$Education,'Education',5)

```

* The distribution of Education is left skewed.
* Majority of customers that purchased the thread mills have atleast 15 years 
of education.


**5. Observations on Usage**
```{r}
plot_histogram_n_boxplot(Fitness$Usage,'Usage',1)

```

*  The distribution of Usage is right skewed.
*  Majority of customers that purchased the thread mills plan on using the 
threadmill aleast 3 times.


**6. Observations on Income**
```{r}
plot_histogram_n_boxplot(Fitness$Income,'Income',2000)

```

*  The distribution of Income is right skewed.
*  Majority of customers that purchased the thread mills have an income of 
 atleast 45000


**7. Observations on Fitness**
```{r}
plot_histogram_n_boxplot(Fitness$Fitness,'Fitness',1)

```

*  The distribution of Income is right skewed.
*  Majority of customers think they have a normal fitness level.


### 3.4 Bivariate Analysis

**1. Observation based on product and gender**
```{r}
Customers_By_Product_Gender = Fitness %>% count(Product, Gender)
          

ggplot(Customers_By_Product_Gender, aes(factor(Product), n, fill = Gender)) +
  geom_bar(stat="identity", position = "dodge") + 
  scale_fill_brewer(palette = "Set1")+
  labs(x = "Gender", 
       y = "Number of Customers", 
       title  = "Number of Customers By Gender and Product")
```

* Checking the distribution of Male and Female customers by Product Type.
* TM798 seems to be more popular amoung Male vs Female Gender


**2. Checking corelation amoung numeric variables**
```{r}
# Checking corelation amoung numeric variables

# Numeric variables in the data
num_vars = sapply(Fitness, is.numeric)

```

```{r}
# Correlation Plot
corrplot(cor(Fitness[,num_vars]), method = 'number')

```

* Usage and Fitness are correlated
* Miles and Fitness are correlated
* Income and Education are correlated


**3. Relationship between Fitness and age, education,usage,income and miles**
```{r}

# Function to draw scatter plots of numerical variables against Product

plot_scatterplot = function(variable, variableNameString, binw){
  ggplot(data = Fitness, aes(x= variable, y = Fitness))+
    labs(x = variableNameString,y ='Fitness')+
    geom_point(color="light blue", size =2, alpha=.8)+
    geom_smooth(method ="lm") # adds a linear trend line which is useful to summarize the relationship between the two variables
}

grid.arrange(plot_scatterplot(Fitness$Age, 'Age'),
             plot_scatterplot(Fitness$Education, 'Education'),
             plot_scatterplot(Fitness$Usage, 'Usage'),
             plot_scatterplot(Fitness$Income, 'Income'),
             plot_scatterplot(Fitness$Miles, 'Miles'),
             ncol = 2)
```

* There seems to be a strong relationship between fitness and usage, fitness 
  and miles.
  

**4. Relationship between Products and age,education,usage,income,miles and 
   fitness**
```{r}
# Function to draw Beeswarm Plot

plot_beeswarm = function(variable, variableNameString, binw){
  ggplot(Fitness, 
       aes(x = factor(Product,
                      labels = c("TM195", #labelling the axis with categorical variable
                                 "TM498", 
                                 "TM798")), 
           y = variable,  
           color = variable)) +
    labs(x = 'Product Code',y = variableNameString, color = variableNameString)+
  geom_quasirandom(alpha = 0.7, 
                   size = 1.5 )  #setting size of dots
}
 

grid.arrange(plot_beeswarm(Fitness$Age, 'Age'),
             plot_beeswarm(Fitness$Education, 'Education'),
             plot_beeswarm(Fitness$Usage, 'Usage'),
             plot_beeswarm(Fitness$Income, 'Income'),
             plot_beeswarm(Fitness$Miles, 'Miles'),
             plot_beeswarm(Fitness$Fitness, 'Fitness'),
             ncol=2)


```
 
 
 * Product TM798 tends to be the preferred model for customers with an income 
   over $80,000
 * TM 195, TM498 seems to be the most popular for users that want to use the 
   product for 3 times or less
 * From the above analysis it is safe to tell that for customers looking to run 
   100 miles or less the   products to go to are TM 195, TM498. 

```{r, include=FALSE}

means<-aggregate(Fitness,by=list(Fitness$Product),mean)

Fitness_group = means[,c("Group.1" ,
          "Age","Education","Usage","Fitness")] ##,"Miles")]
#View(Fitness_group)

means.long<-  melt(Fitness_group,id.vars="Group.1")

```

  

**5. Compare mean of age, education, usage,fitness across the 3 products **
```{r}
box= ggplot(means.long,aes(x=variable,y=value,fill=Group.1))+
        geom_bar(stat="identity", position=position_dodge())+
        scale_fill_discrete(name="Product",
                      breaks=c("T195", "T498","T798"),
                      labels=c("T195", "T498","T798"))+
  xlab("Product")+ylab("Mean Percentage") 
box

```


* Product TM798 seems to be the product that people seem to choose if they think 
  they have a high fitness and are going to use the product more than an average 
  customer


### 4.Conclusion and Recommendation

#### 4.1 Conclusion

Based on the analysis of the Cardio Good Fitness data set below are the insights 
that we have been able to deduce.

* Products T195 seems to be the most popular treadmill out of the 3.
* Majority of the customers seem to be in their 20's or early 30's,
  basically younger people.
* There is a significantly higher number of Male customers when compared to 
  female customers.
* A significant number of customers make less than $60,000.
* Majority of the customers would like to use the treadmills for 3 times a week.
* The average number of miles majority of the customers purchasing the T195, 
  T498 treadmill want to run or walk is less than 100 per week.
* Product T798 is purchased by people that want to use the treadmill more than 
  the average usage and usually their fitness levels are higher than the average 
  fitness among all the customers from the dataset.
  
<br /> 
<br /> 
<br /> 
<br /> 
<br /> 
<br /> 

#### 4.2 Recommendation
 From the observations made from the data there are 2 distinctive customer 
 profiles and below are the recommendations.


**Targeted marketing campaigns for consumers with below characteristics**

* By doing a targeted market campaign the company can effectively use the 
  allocated marketing funds.

**Consumer Profile I**


* Consumers that use Product T195 and T498 usually use their treadmills for 
  3 times a week and approximately want to run less than or equal to 100 miles
  a week.
* These products seem to be more budget friendly since all the customers that 
  purchased these treadmills have an income of less than 70,000.
* Products  T195 and T498 can be marketed for the consumers with these 
  characteristics.

<br /> 
<br /> 

**Consumer Profile II**


* Consumers purchasing the T798 seem to be people that exercise more than an 
  average customer and are willing to spend the money on it.
* All the customers that purchased T798 either make more than 70,000 or have a
  usage higher than the average customer from the group.
* Product T798 seems to be a luxurious model and is preferred by fitness 
  enthusiasts.
* Product T798 can be marketed for consumers based on the above characteristics.  
