---
title: "CardioFit Analysis"
layout: post
date: "2020-07-10 23:00:40 +0100"
categories: r jekyll
---

  
## 1. Project Objective
  
  The objective of this report is to explore the Cardio Good Fitness data set in 
R and extract insights about the data.

This exploration report will consist of the following:
  
* Importing the dataset in R 
* Understanding the structure of dataset
* Graphical Exploration
* Descriptive Statistics
* Generate insights and recommendations that will help the company in targeting 
new customers

## 2. Assumptions 

The dataset provided is a sample of the population for customers that bought the 
treadmill product of Cardio Good Fitness Store. 

## 3. Exploratory Data Analysis - Step by step approach

```{r echo = FALSE}
#======================================================================= 
# 
# Exploratory Data Analysis - CardioGoodFitness 
# 
#=======================================================================
```

### 3.1 Environment Set up and Data Import 

#### 3.1.1 Install necessary packages and load libraries
```{r,warning=FALSE, message=FALSE}


library(ggplot2)# To create plots
library(gridExtra)# To plot multiple ggplot graphs in a grid
library(dplyr) # For data manipulation
library(knitr)# Necessary to generate sourcecodes from a .Rmd File 
library(corrplot) # To plot correlation plot between numerical variables
library(reshape2)# To easily transform data between wide and long formats
library(plotly) # R package for creating interactive web-based graphs 
library(ggbeeswarm) #R package for plotting categorical scatter plots
library(scales) # R package for internal scaling infrastructure used by ggplot2 
library(webshot)

```

#### 3.1.2 Set up working Directory

```{r}
# Setup Working Directory 
# Rmd considers it's directory as the working directory.
# If the Rmd and data file is in the same directory there is no need of 
# setting up the working directory.

# While working with .R files,we need to set up the working directory using the following command
# setwd("C:/Users")

```


#### 3.1.3 Import and Read the Dataset

```{r}
# Read Input File 
Fitness = read.csv("CardioGoodFitness.csv")
```


#### 3.1.4 Global options settings
```{r}
# Global options settings 
options(scipen=999) # turn off scientific notation like 1e+06
```

### 3.2 Variable Identification

#### 3.2.1 Variable Insights

**1. Checking the dimension of the dataset**
```{r}
#dim(Fitness)
knitr::kable(head(Fitness,10))

```
* The dataset has 180 rows and 9 columns


```

